Efficient UQ for chemical systems using active \author{Manav Vohra, Alen Alexanderian, Hayley Guy, Sankaran Mahadevan} subspaces

### 2018-09-24 Paper/abstract.tex
abstract ...

\section*{Abstract}
Time evolution of species' concentration in a chemical system is tightly coupled
with the underlying mechanism and the governing rate laws for individual reactions.
Oftentimes, the rate-controlling parameters cannot be specified precisely 
due to sparse calibration data and lack of knowledge about them, regarded as
the \textit{epistemic} uncertainty. Computer simulations for such systems must
therefore account for the presence of the epistemic uncertainty in these parameters
when making predictions. However, complex mechanisms typically involve a large number of
reactions and consequently, a large set of uncertain rate parameters. Propagation of
the uncertainty from parameters to predictions can be remarkably challenging for
such high-dimensional problems. In this study, we focus on enabling an efficient
approach for uncertainty propagation from inputs to the model output,
through dimension reduction by computing the
so-called
\textit{active subspace}. The active subspace
 predominantly captures the variability in the quantity of
interest (QoI) in terms of fewer independent variables. 
It is computed using an iterative, gradient-based
approach and a computationally less intensive gradient-free approach for the 
H$_2$/O$_2$ mechanism involving 19 reactions. An initial investigation considers
only the 19 pre-exponents of the Arrhenius rate laws as uncertain. In this case, both
approaches yield consistent results wherein the active subspace is found to be 
1-dimensional. The two approaches are further implemented to a high-dimensional
problem involving 33 uncertain rate-parameters: pre-exponent and the activation
energy. Once again, a 1-dimensional active subspace is obtained, indicating
enormous scope for dimension reduction in such applications. However, the
gradient-free approach though reasonably captured the mean, mode, and the uncertainty
in the QoI; failed to capture the sensitivity towards the most
important parameter according to the gradient-based analysis for the high-dimensional
application.

>>>>

\section*{Abstract}
We focus on efficient quantification of uncertainty in complex chemical
reaction networks with a large number of uncertain parameters.  Parameter
dimension reduction is accomplished by computing an~\emph{active subspace} that
predominantly captures the variability in the quantity of interest (QoI).  In
this study, we compute the active subspace for the H$_2$/O$_2$ mechanism that
involves 19 chemical reactions, using an efficient iterative strategy.  The
active subspace is first computed for a 19-parameter problem wherein only the
uncertainty in the pre-exponents of the individual reaction rates is
considered. This is followed by the analysis of a 33-dimensional case wherein
the activation energies are also considered uncertain.  In both cases, a
one-dimensional active subspace is identified, which indicates
enormous potential for efficient statistical analysis of complex 
chemical systems. In addition, we explore links between active subspaces and
global sensitivity analysis measures, and exploit these links for 
indentification of key contributors to the variability in the model response.

%In this case, the two approaches are observed to yield consistent results.
%However, when computing the active subspace for a 33-dimensional problem
%involving uncertain activation energies in addition to the pre-exponents, a
%trade-off between accuracy and effort is observed.

%Specifically, when estimating the gradient using
%the regression-based approach, the variability of the QoI (in the subspace),
%due to uncertainty in the rate parameters, is under-estimated. Moreover, the
%sensitivity towards one of the most important parameter is not captured. Hence,
%we carefully recommend using the regression-based approach for estimating the
%gradient only in situations involving intensive computations of the model 
%(e.g. complex mechanisms involving a large number of reactions and species), and where the
%goal is to obtain rough estimates of the statistics of the QoI.
